---
title: "Credit Card Defaulter Prediction Model Analysis"
author: "Eric Fruelda"
date: "Thursday, January 28, 2016"
output: html_document
---

##About this Analysis

This analysis uses a dataset of 400 cases of credit card customers with default and non-default outcomes. This document provides the detailed procedure in the creation of a prediction model using multivariate linear regression for predicting future defaulters and the corresponding analysis for the model.

##Dataset

The dataset contains 400 credit card customers with 100 defaulters and 300 non-defaulters. There are six pieces of information that were provided for each customer:

+ Age
+ Years at current employer
+ Years at current address
+ Income over the past year
+ Current credit card debt
+ Current automobile debt.

The first column corresponds to the unique identifier "Unique.ID" for each customer and the last column corresponds to the outcome; 1 for default and 0 for non-default.

##Methodology

The following steps were done for this project:

1. Data Processing
    + Load the data.
    + Clean the data.
    + Split the data.
2. Building the Model
    + Fit a multivariate linear regression model.
    + Test the model.

##STEP 1. Data Processing

###Loading Data
Loading libraries and dataset.
```{r loaddata, echo=TRUE, results='hide'}
library(caret )
data <- read.csv("data.csv", header = TRUE)
```

###Data Cleaning
Selecting useful variables by leaving out the "Unique.ID" column.
```{r cleandata, echo=TRUE, results='hide'}
data<-data[,c(2:8)]
```

###Splitting Data
Spliting the training data to 75% training set using the first 300 cases and 25% testing set using the remaining 100 cases.
```{r splitdata, echo=TRUE, results='hide'}
training <- data[c(1:300), ] 
testing <- data[c(301:400), ]
```

##STEP 2. Building the Model

###Fitting Multivariate Linear Regression Model
Using the lm function to fit multivariate linear regression model.
```{r fitLM, echo=FALSE}
fitmodel <- lm(Default_Outcomes~.,training)
print(fitmodel) # Displaying model
```

###Testing Model
The resulting model is evaluated using the testing set splitted in the beginning. A confusion matrix is provided with the testing set to evaluate accuracy of the model.
```{r testLM, echo=FALSE}
confusionMatrix(testing$Default_Outcomes,round(predict(fitmodel,testing)))
```
The model had an overall accuracy of 0.83, or 83% at 95% CI:(0.7418, 0.8977). Therefore, the out-of-sample error rate was 0.17, or 17%.

## Discussions
In this analysis, 75% of the data were used to build a model by multivariate linear regression with the remaining 25% used for model testing. The model statistics by class showed that the built model had the overall accuracy of 83% for the testing set. The out-of-sample error rate was 17%. The sensitivity was 82.61% and the specificity was 87.5%. With an accuracy above 83% on our cross-validation data, we can expect that 4 out of 5 customers can be accurately predicted whether to be a future defaulter or not.

## Source of Data and Further Reading
For further reading, please read the document "Information on Credit Card Customer Data" that comes along with this analysis. This document along with the data file is available in [Github](https://github.com/ericfruelda/Developing-Data-Products.git).